### 2021-08-18 manuscript/SI.tex

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames*,x11names*}{xcolor}
%
\documentclass[
  11pt,
]{article}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{setspace}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={Supplementary Data for Testing the motor and cognitive foundations of Paleolithic social transmission},
  pdfauthor={Justin Pargeter; Megan Beney Kilgore; Cheng Liu; Aditi Majoe; Dietrich Stout},
  colorlinks=true,
  linkcolor=blue,
  filecolor=Maroon,
  citecolor=Blue,
  urlcolor=Blue,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage[margin=1in]{geometry}
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
\usepackage{textcomp}
\usepackage{utopia}
\usepackage{lineno}
\linenumbers
\usepackage{float}
\floatplacement{figure}{H}
\usepackage{caption}
\captionsetup[figure]{font=scriptsize}
\captionsetup[table]{font=scriptsize}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\ifluatex
  \usepackage{selnolig}  % disable illegal ligatures
\fi

\title{Supplementary Data for Testing the motor and cognitive foundations of Paleolithic social transmission}
\author{Justin Pargeter\footnote{Department of Anthropology, New York University, New York, NY, USA; Palaeo-Research Institute, University of Johannesburg, Auckland Park, South Africa; \href{mailto:justin.pargeter@nyu.edu}{\nolinkurl{justin.pargeter@nyu.edu}}} \and Megan Beney Kilgore\footnote{Department of Anthropology, Emory University, Atlanta, GA, USA; \href{mailto:megan.elizabeth.beney@emory.edu}{\nolinkurl{megan.elizabeth.beney@emory.edu}}} \and Cheng Liu\footnote{Department of Anthropology, Emory University, Atlanta, GA, USA; \href{mailto:raylc1996@outlook.com}{\nolinkurl{raylc1996@outlook.com}}} \and Aditi Majoe\footnote{Indepedent Researcher; \href{mailto:am2752@cantab.ac.uk}{\nolinkurl{am2752@cantab.ac.uk}}} \and Dietrich Stout\footnote{Department of Anthropology, Emory University, Atlanta, GA, USA; \href{mailto:dwstout@emory.edu}{\nolinkurl{dwstout@emory.edu}}}}
\date{}

\begin{document}
\maketitle

{
\hypersetup{linkcolor=}
\setcounter{tocdepth}{2}
\tableofcontents
}
\setstretch{1.5}
\begin{verbatim}
## Warning in read.table(file = file, header = header, sep = sep, quote = quote, :
## invalid input found on input connection 'C:/FilesVC/PaST-pilot/data/tables/
## s1.csv'
\end{verbatim}

\begin{verbatim}
## Warning in read.table(file = file, header = header, sep = sep, quote = quote, :
## line 1 appears to contain embedded nulls
\end{verbatim}

\begin{verbatim}
## Warning in read.table(file = file, header = header, sep = sep, quote = quote, :
## incomplete final line found by readTableHeader on 'C:/FilesVC/PaST-pilot/data/
## tables/s1.csv'
\end{verbatim}

\begin{table}[H]

\caption{\label{tab:tab1}Lithic performance between two groups.}
\centering
\fontsize{7}{9}\selectfont
\begin{tabular}[t]{l}
\toprule
\cellcolor{gray!6}{PK...}\\

\bottomrule
\end{tabular}
\end{table}

\begin{verbatim}
## Warning in read.table(file = file, header = header, sep = sep, quote = quote, :
## invalid input found on input connection 'C:/FilesVC/PaST-pilot/data/tables/
## s2.csv'
\end{verbatim}

\begin{verbatim}
## Warning in read.table(file = file, header = header, sep = sep, quote = quote, :
## line 1 appears to contain embedded nulls
\end{verbatim}

\begin{verbatim}
## Warning in read.table(file = file, header = header, sep = sep, quote = quote, :
## incomplete final line found by readTableHeader on 'C:/FilesVC/PaST-pilot/data/
## tables/s2.csv'
\end{verbatim}

\begin{table}[H]

\caption{\label{tab:tab2}Flake metrics between two groups.}
\centering
\fontsize{7}{9}\selectfont
\begin{tabular}[t]{l}
\toprule
\cellcolor{gray!6}{PK...}\\

\bottomrule
\end{tabular}
\end{table}

\begin{verbatim}
## Warning in read.table(file = file, header = header, sep = sep, quote = quote, :
## invalid input found on input connection 'C:/FilesVC/PaST-pilot/data/tables/
## s3.csv'
\end{verbatim}

\begin{verbatim}
## Warning in read.table(file = file, header = header, sep = sep, quote = quote, :
## line 1 appears to contain embedded nulls
\end{verbatim}

\begin{verbatim}
## Warning in read.table(file = file, header = header, sep = sep, quote = quote, :
## incomplete final line found by readTableHeader on 'C:/FilesVC/PaST-pilot/data/
## tables/s3.csv'
\end{verbatim}

\begin{table}[H]

\caption{\label{tab:tab3}Starting nodule metrics between two groups.}
\centering
\fontsize{7}{9}\selectfont
\begin{tabular}[t]{l}
\toprule
\cellcolor{gray!6}{PK...}\\

\bottomrule
\end{tabular}
\end{table}

\end{document}