### 2021-08-21 manuscript/manuscript.Rmd
updated the main text and supplementary tables.

---
title: Testing the social, cognitive, and motor foundations of Paleolithic skill reproduction
author: 
- Justin Pargeter^[Department of Anthropology, New York University, New York, NY, USA; Palaeo-Research Institute, University of Johannesburg, Auckland Park, South Africa; justin.pargeter@nyu.edu]
- Cheng Liu^[Department of Anthropology, Emory University, Atlanta, GA, USA; raylc1996@outlook.com]
- Megan Beney Kilgore^[Department of Anthropology, Emory University, Atlanta, GA, USA; megan.elizabeth.beney@emory.edu]
- Aditi Majoe^[Indepedent Researcher; am2752@cantab.ac.uk]
- Dietrich Stout^[Department of Anthropology, Emory University, Atlanta, GA, USA; dwstout@emory.edu]
abstract: "Stone tools provide key evidence of human cognitive evolution but remain challenging to interpret. Tool-making skill-learning has been understudied even though: 1) the most salient cognitive demands of tool-making should occur during learning, and 2) variation in learning aptitude would have provided the raw material for any past selection acting on tool-making ability. However, we know very little about the cognitive prerequisites of learning under different information transmission conditions that may have prevailed during the Paleolithic. This paper presents results from an exploratory study to trial new experimental methods for studying the effect of learning conditions and individual differences on Oldowan flake-tool-making skill acquisition. We trained 23 participants for 2 hours to make stone flakes under two different instructional conditions (observation only vs. direct active teaching). We employed appropriate raw materials, a moderate practice time duration, and in-person, fully interactive instruction. Participant performance was evaluated through an analysis of the stone artifacts produced. We compared performance across experimental groups with respect to individual participant differences in grip strength, motor accuracy, and cognitive function measured for the study. Our results show that 2 hours are insufficient to document learning-related performance change. However, teaching reduces variability in knapping rate, methods, and outcomes during early-stage learning, thus increasing the reliability of skill reproduction. Instruction also altered knapping quality vs. quantity trade-offs in the two groups and dramatically changed the effects of individual differences in strength, visuospatial working memory, and social learning tendencies on knapping outcomes. Our results provide further support for the hypothetical co-evolution of teaching, language, and tool-making, suggest that the presence/absence of instruction can fundamentally alter learning-related selection pressures on individuals, and offer lessons for the design of future experiments.\\par \\textbf{Keywords:} Oldowan; Stone tool-making; Social learning; Individual variation; Cognitive aptitudes; Motor skills"
fontsize: 11pt
output: bookdown::pdf_document2
header-includes:
   - \usepackage{textcomp}
   - \usepackage{utopia}
   - \usepackage{lineno}
   - \linenumbers
   - \usepackage{float}
   - \floatplacement{figure}{H}
   - \usepackage{caption}
   - \captionsetup[figure]{font=scriptsize}
   - \captionsetup[table]{font=scriptsize}
linestretch: 1.5
link-citations: yes
linkcolor: blue
bibliography: bibliography.bib
csl: apa.csl
---

# Introduction

Stone tools have long been seen as a key source of evidence for understanding human behavioral and cognitive evolution [@darwin1871; @oakley1949; @washburn1960]. Pathbreaking attempts to infer specific cognitive capacities from this evidence largely focused on the basic requirements of tool production [@isaac1976; @wynn1979; @gowlett1984; @wynn2004]. More recently, researchers have directed increasing attention to the processes and demands of stone tool-making skill acquisition [@roux1995; @stout2002; @stout2005; @geribàs2010; @nonaka2010; @stout2011; @putt2014; @hecht2015a; @duke2015; @morgan2015; @stout2015; @lombao2017; @putt2017; @cataldo2018; @putt2019; @pargeter2019; @pargeter2020; @shipton2018]. This increased attention is motivated by the expectation that tool-making's most salient cognitive demands should occur during learning rather than routine expert performance [@stout2015]. Researchers are also increasingly focused on the relevance of different social learning mechanisms such as imitation [@rein2014; @stout2019], emulation [@tehrani2008; @wilkins2018], and language [@ohnuma1997; @putt2014; @morgan2015; @lombao2017; @putt2017; @cataldo2018] to the reproduction of Paleolithic technologies.

Studies investigating these questions have used a range of different experimental designs (e.g., varying technological goals/instructions, training times, raw materials, live vs. recorded instruction, lithic/skill assessment metrics, pseudo-knapping tasks etc.) and reached disparate conclusions regarding skill acquisition’s neurocognitive and social foundations. It is plausible that these discordant results reflect actual diversity in how humans acquire and master stone tool-making skills. However, this failure of results to generalize across artificial experimental manipulations [cf. @yarkoni2020] also raises doubts regarding the external validity [@eren2016] of conclusions concerning real-world Paleolithic learning contexts. To address this, we conducted an exploratory study that draws on lessons from previous research to balance the pragmatic and theoretical trade-offs inherent in experimental studies of stone knapping skill acquisition [@pargeter2019; @stout2015].

Compared to artificial experimental tasks, learning real-world skills like stone knapping is highly demanding of time and materials and difficult to control experimentally without sacrificing generalizability to real world conditions. Prior efforts have attempted to navigate these challenges by using various combinations of 1) inauthentic raw materials that are less expensive, easier to standardize, and/or easier to knap, 2) video-recorded instruction that is uniform across participants and less demanding of experimenter time, 3) short learning periods, 4) small sample sizes, and 5) single learning conditions. The difficulty of interpreting results from this growing literature led Stout and Khreisheh [-@stout2015: 870, emphasis original] to call for "studies with sufficient sample sizes to manipulate learning conditions (e.g. instruction, motivation) and assess individual variation (e.g. performance, psychometrics, neuroanatomy) that *also* have realistic learning periods." The current study attempts to strike a viable balance between these demands by investigating early-stage learning of a relatively simple technology (least effort, "Oldowan," flake production [@reti2016; @shea2016]) under two instructional conditions while collecting data on individual differences in strength, coordination, cognition, and social learning. Unlike any previous study, this allows us to address the likelihood that interactions with individual participant differences in aptitude or learning style might impact group effects of training conditions.

We focus on early-stage learning because it is relatively rapid, variable across individuals, and predictive of later outcomes [@pargeter2019; @stout2015; @putt2019]. The study thus provides a reasonable expectation of generating meaningful data on skill and learning variation while minimizing training costs. Moreover, understanding the minimum training times necessary to detect changes in tool making skill will help archaeologists design more realistic and cost-effective experiments. We limited our study to only two learning conditions (observation only vs. active teaching) to further manage costs. This study design targets a key controversy in human evolution, namely the origins of teaching and language [@gärdenfors2017; @morgan2015] while avoiding artificial manipulations of dubious relevance to real-world Paleolithic learning [e.g., @cataldo2018]. These choices allowed us to invest more in other aspects of research design that we identified as theoretically necessary, including measuring individual differences in cognition and behavior, including an in-person, fully interactive teaching condition, and using naturalistic raw materials. The sample size remained small in this internally funded exploratory study, but we could easily scale it up at funding levels typical of pre- and post-doctoral research grants in archaeology.

## Individual Differences

|   "*The many slight differences... being observed in the individuals of the same species inhabiting the same confined locality, may be called individual differences... These individual differences are of the highest importance to us, for they are often inherited ... and they thus afford materials for natural selection to act on and accumulate...*" [@darwin1859, Chapter 2]

Individuals vary in aptitude and learning style for skills [@jonassen1993], but research on knapping skill acquisition has largely ignored this. These studies have instead focused on group effects of different experimental conditions. There are good pragmatic reasons for this, as individual difference studies typically require larger sample sizes and other data collection. However, overlooking these distinctions is not ideal since individual differences can provide valuable insight into cognition and behavior's mechanisms, development, and evolution [@boogert2018]. Researchers can use patterns of association between cognitive traits and behavioral performance to test hypotheses about the cognitive demands of learning particular skills and the likely targets of natural selection acting on aptitude. More prosaically, individual differences can introduce an unexamined and uncontrolled source of variation in group-level results. This effect is especially true in the relatively small samples of convenience typical of experimental archaeology.

While testing hypotheses in evolutionary cognitive archaeology remains a considerable challenge [@wynn2017], investigating individual variation in modern research participants represents one promising direction. For any behavior of archaeological interest, we could expect that standing variation in modern populations should remain relevant to normal variation in learning aptitude. The presence of trait variation without impact on learning aptitude would provide strong evidence against the plausibility of the proposed evolutionary relationship. We would not expect an an absence of variation (i.e., past fixation and rigorous developmental canalization) given the known variability of human brains and cognition [@sherwood2017; @barrett2020]. Any confirmatory findings of trait-aptitude correspondence would have the testable implication that humans should be evolutionarily derived along the same dimension [@hecht2015b].

To date, a small number of "neuroarchaeological" studies have reported associations between individual knapping performance and brain structure or physiological responses. Hecht et al. [-@hecht2015a] reported training-related changes in white matter structure correlated with individual differences in practice time and striking accuracy change. The regional patterning of white matter changes also varied across individuals. Only those who displayed early increases under the right ventral precentral gyrus (premotor cortex involved in movement planning and guidance) showed striking accuracy improvement over the training period. Putt et al. [-@putt2019] similarly found that the proportion of flakes-to-shatter produced by individuals during handaxe-making correlated with dorsal precentral gyrus (motor cortex) activation. Pargeter et al. [-@pargeter2020] used a flake prediction paradigm [modeled after @nonaka2010] to confirm that striking force and accuracy are important determinants of handaxe-making success. These findings point to the central role of perceptual-motor systems [@stout2007] and coordination [@roux1995] in knapping skill acquisition. In addition, Putt et al. [-@putt2019] found successful flake production to be associated with prefrontal (working memory/cognitive control) activation, and Stout et al. [-@stoutCognitiveDemandsLower2015] found that prefrontal activation correlated with success at a strategic judgement (platform selection) task which in turn was predictive of success at out-of-scanner handaxe production.

Such investigations are thus starting to chart the contributions of different neural systems to aspects of knapping skill acquisition, including motor and cognitive control aspects. To date, however, the cognitive/functional interpretation of systems identified in this manner has relied mainly on informal reverse inference (reasoning backward from observed activations to inferred mental processes) from published studies of other tasks that activated the same regions. Researchers widely regard this approach as problematic [@poldrack2011]. Here we take a more direct, psychometric approach to measuring individual differences in perceptual-motor coordination and cognition.

Studies implement psychometric instruments (e.g., tasks, questionnaires) to assess variation in cognitive traits and states, such as fluid intelligence, working memory, attention, motivation, and personality. These measures have been of theoretical interest to the broader field of cognitive archaeology [e.g., @wynn2016]. It is thus surprising that experimental studies have almost entirely neglected them. In the only published example we are aware of, Pargeter et al. [-@pargeter2019] reported significant effects of variation in planning and problem solving (Tower of London test [@shallice1982]) and cognitive set shifting (Wisconsin Card Sort test [@grant1948]) on early stage handaxe learning. Of course, cognition is not the only thing that can affect knapping performance. Flake prediction experiments highlight the importance of regulating movement speed/accuracy trade-offs [@nonaka2010; @pargeter2020] and studies of muscle recruitment [@marzke1998] and manual pressure [@williams-hatala2018; @key2018] during knapping highlight basic strength requirements. Along these lines, Key and Lycett [-@key2019] found that individual differences in hand size, shape, and especially grip strength were better predictors of force loading during stone tool use than were attributes of the tools themselves. However, we are unaware of any such studies of biometric influences on variation in knapping success. Finally, individual, and contextual differences in adopting social vs. individual learning strategies [@miu2020] might also affect learning outcomes. We are again unaware of any previous studies that have assessed such effects. This study evaluated all participants with a battery of tests, including grip strength, movement speed/accuracy, spatial working memory, fluid intelligence, and tendency to use social information. We were particularly interested in the possibility that these variables might not only impact learning generally but might also have different effects under different learning conditions.

## Teaching, Language, and Tool Making

|   "*A creature that learns to make tools to a complex pre-existing pattern...must have the kind of abstracting mind that would be of high selective value in facilitating the development of the ability to communicate such skills by the necessary verbal acts.*"[@montagu1976: 267]

Possible links between tool making and language have been a subject of speculation for nearly 150 years [@engles2003[1873]], if not longer [@hewes1993], although compelling empirical tests have remained elusive. Over 25 years ago, Toth and Schick [-@toth1993] suggested that experiments teaching modern participants to make stone tools in verbal and non-verbal conditions could test the importance of language in the social reproduction of Paleolithic technologies. Ohnuma et al. [-@ohnuma1997] were the first to implement this suggestion in a study of Levallois flake production, followed by more recent studies of handaxe making [@putt2014; @putt2017] and simple flake production [@morgan2015; @cataldo2018; @lombao2017]. This study design trajectory reflects a recent interest in the hypothesis that language might be an adaptation for teaching [e.g., @laland2017; @stout2012]. Teaching and learning demands of Paleolithic tool-making would thus provide evidence of selective contexts favoring language evolution [@stout2010; @morgan2015; @montagu1976].

However, Toth and Schick [-@toth1993] pointed out that extinct hominid learning strategies and capacities might differ from modern experimental participants. Even leaving aside potential species differences in social learning [cf. @morgan2015; @stout2019], reliance on explicit verbal instruction varies widely across modern human societies [e.g., @boyette2017]. The WEIRD (Western, educated, industrialized, rich, democratic [@henrich2010]) teachers and learners typical of knapping experiments arguably represent an extreme bias toward such instruction. Simply instructing such participants not to speak during an experiment (or to demonstrate but not gesture, etc. [@morgan2015]) is likely to underestimate the efficacy of non-verbal teaching and learning in cultural contexts where it is more common, let alone in a hypothetical pre-linguistic hominid species [cf. @fuentes2015].

Such concerns are exacerbated in experiments using pre-recorded instructional videos or brief training periods. Video does not allow the interactive teaching favored even in formal academic knapping classes [e.g., @shea2015] and is almost undoubtedly typical of traditional learning contexts [e.g., @stout2002]. Researchers do not know how video presentation generally affects the efficacy of teaching stone knapping or the relative effectiveness of different forms of instruction. Moreover, some experiments have manipulated the presence/absence of verbal instruction by presenting the same video with and without sound [@putt2017] or the sound track without the video [@cataldo2018]. While this provides experimental control, it does not allow the instructor to adjust their multimodal [@levinson2014] communication strategies as they would naturally do, for example, through pointing and pantomime. Simply removing a communication channel without allowing such adaptation is highly artificial and risks generating results that cannot be generalized beyond the specific context of the experiment [@yarkoni2020]. Similarly, unnaturally short training periods (e.g., 5-15 minutes [@morgan2015; @lombao2017]) might misrepresent the relative efficacy of different teaching strategies under more realistic conditions [@whiten2015; @stout2015]. Even the most extended training times [@stout2015; @pargeter2019] have not produced knapping skills comparable to relevant archaeological examples and were achieved by limiting sample size and using only one teaching condition.

For these reasons, we sought to explore a middle path between experimental expedience and realism by limiting our experiment to two relatively naturalistic learning conditions and a moderate learning period of two hours. As in previous experiments [@stout2011; @hecht2015a; @pargeter2019], the first condition was unrestricted, interactive instruction in small groups, essentially reproducing the "natural" teaching/learning context familiar [cf. @shea2015] to our WEIRD instructor and student participants. The second condition allowed observation only, with the experimenter visible making flakes but not providing instruction or interacting with learners in any way during this activity. This absence of teaching is again a familiar social context for our participants and did not require any novel behaviors from the instructor. Although we make no assumptions regarding learning mechanisms, it matches the "imitation/emulation" condition of Morgan et al. [-@morgan2015]. We did not include a "reverse engineering" or "end-state emulation" condition where only finished products were visible. Researchers have advocated these states as an important baseline or control condition [@whiten2015] to distinguish observational from individual learning. However, they are not likely to model any typical Paleolithic learning context nor to stand as an adequate proxy for the cognition of hominid species with different social learning capacities. There is no reason to assume neurocognitive and behavioral processes of reverse-engineering problem solving in modern humans [e.g., @allen2020] approximate the social learning processes of hominids with more ape-like action observation/imitation capacities [@hecht2013a; @hecht2013b; @stout2019].

We selected a two-hour learning period for both pragmatic and theoretical reasons. Pargeter et al. [@pargeter2019] found that even \~90 hours of fully interactive instruction and practice was insufficient to achieve handaxe-making skills comparable to the later Acheulean site of Boxgrove [@garcía-medrano2019; @stout2014], and estimated actual time to mastery as ranging from 121 to 441 hours for different participants. However, they observed the greatest, fastest, and most individually variable skill increases during the first 20 hours of practice. In addition, initial performance moderately correlated with later achievement. This result suggests that studying early-stage learning may be a practical alternative, especially for research investigating individual differences in aptitude. Studies of simple flake production similarly document considerable initial variation [@stout2015] and rapid early progress [@stout2015; @putt2019; @stoutKnappingSkillEarliest2006]. We designed the current study to test the utility of studying learning and variation during the first two hours of simple flaking instruction/practice in hopes of finding a viable compromise between experimental realism and cost.

## Raw materials and knapping skill

|   "*Such undertakings -- based on raw material which is never standard, and with gestures of percussion that are never perfectly delivered -- cannot be reduced to an elementary repetition of gestures...the realization of elaborate knapping activities necessitates a critical monitoring of the situation and of the decisions adopted all through the process." [@pelegrin1990: 117]*

Lithic raw materials vary in size, shape, and fracture mechanical properties that affect the difficulty of achieving different knapping goals [@eren2014]. Unfortunately, it can be difficult and/or expensive to procure authentic raw materials. Experimental studies of knapping skill have often used proxy materials such as flint [@morgan2015; @nonaka2010; @cataldo2018], limestone [@stoutKnappingSkillEarliest2006], porcelain [@khreisheh2013], or heat-treated chert [@putt2014; @putt2017; @putt2019] to model Oldowan and early Acheulean technologies executed in other materials. As well as being more readily available, these proxies are generally easier to knap. This study design has the benefit of reducing required practice time. Still, it is unclear how it might affect learning demands more generally or the efficacy of different learning conditions/strategies precisely.

Some designs have attempted to closely match experimental and archaeological raw material types [@stout2011; @duke2015; @pargeter2019]. However, raw materials vary across individual clasts within as well as between types. This choice has led to an interest in standardizing experimental core morphology [@nonaka2010] and composition, even if this means using artificial materials such as porcelain [@khreisheh2013], brick [@geribàs2010; @lombao2017], or foam blocks [@schillinger2014]. Such manipulations enhance experimental control and internal validity [@eren2016] at the expense of external generalizability to actual archaeological conditions. Specifically, they allow more robust results from smaller samples but eliminate a core element of real-world knapping skill: the ability to produce consistent results from variable materials [@pelegrin1990; @stout2013]. For example, Pargeter et al. [-@pargeter2020] found that predicting specific flaking outcomes on actual handaxe preforms was both more complex and less technologically important than expected from previous work with standardized, frustum-shaped cores [@nonaka2010]. The alternative to control is to incorporate raw material size, shape, and composition as experimental variables [e.g., @stout2019]. This choice allows consideration of raw material selection and response to variation as aspects of skill but correspondingly increases the sample sizes required to identify patterning. In considering these issues, we again chose to explore a middle path between pragmatism and realism by employing commercially purchased basalt like that known from East African Oldowan sites. We allowed clast size and shape to vary within set limits and selected the particular clasts provided to each participant to approximate the same distribution.

# Materials and Methods

The Emory Institutional Review Board approved this research (IRB00113024). All participants provided written informed consent and completed a video release form (<https://databrary.org/support/irb/release-template.html>).

## Participants

Twenty-four adult participants with no prior stone knapping experience were recruited from the Emory community using paper fliers and e-mail listserv advertisements. Eleven participants (6 female, five male) completed the uninstructed condition, and 12 (8 female, four male) completed the instructed condition. We could not replace one participant who failed to attend their scheduled session, resulting in a total sample of 23.

## Study Visit

We asked participants to visit the Paleolithic Technology Lab at Emory University to complete one three-hour session. Participants attended in six groups of four, however, one of these groups had only three due to a no-show on the day of the experiment. Each visit began with the collection of individual differences measures, which took approximately one hour. After that, participants undertook 105 minutes (two hours minus a 15-minute break after 1 hour) of stone tool-making practice. This session was video-recorded, and we collected all lithic products. After the tool-making task, participants completed an "exit questionnaire" comprising the Intrinsic Motivation Inventory (see below).

We compensated participants for their time with a \$30 gift card. They also had the opportunity to earn a performance bonus of \$5, \$10, \$15 or \$20 on the gift card. They were told that this bonus would depend on "how well they did" on the last core of their practice session. The actual performance measure was not specified, but to allow on the spot payment we used a simple measure of the percentage of starting weight removed from the final core such that: \> 30% earned \$5, \> 40% earned \$10, \> 50% earned \$15, \> 75% earned \$20.

## Individual Difference Measures

We used five individual difference measures for this study:

1)  *Grip strength* was measured in kilograms using an electronic hand dynamometer (Camry EH101). We measured strength twice and kept the higher value. Grip strength is a simple measure well correlated with overall muscular strength [@wind2010] and a range of other health and fitness measures [@sasaki2007]. Researchers have hypothesized that grip strength is relevant for generating kinetic energy for fracture initiation [@nonaka2010] as well as control and support of the hammerstone [@williams-hatala2018] and core [@faisal2010; @key2015].

2)  *Motor accuracy* was assessed using a "Fitts Law" reciprocal tapping task. Fitts Law describes the trade-off between speed and accuracy in human movement, classically measured by tapping back and forth between two targets of varying size and spacing [@fitts1954]. Archaeologists have proposed [@stout2002; @pargeter2020] that managing this trade-off is critical to the accurate application of appropriate force seen in skilled knapping [@nonaka2010; @roux1995]. We implemented this test on a Surface Pro tablet running free software (FittsStudy Version 4.2.8, default settings) developed by the Accessible Computing Experiences lab (Jacob O. Wobbrock, director) at the University of Washington ([depts.washington.edu/acelab/proj/fittsstudy/index.html](depts.washington.edu/acelab/proj/fittsstudy/index.html)). Participants use a touchscreen pen to tap between ribbons on the screen, with average movement time as the performance metric.

3)  *Visuospatial working memory* is the capacity to "hold in mind", which researchers have hypothesized to be important in stone tool-making performance [@coolidge2005]. It also might support a learning process known as 'chunking', in which subjects combine multiple items or operations into summary chunks stored in long-term memory, which researchers think is essential in acquiring knapping and other skills [@pargeter2019]. We measured visuospatial working memory using a free n-back task ([wmp.education.uci.edu/software/](https://wmp.education.uci.edu/software/)) developed by the Working Memory and Plasticity Laboratory at the University of California, Irvine (Susanne Jaeggi, PI) and implemented in E-Prime software on a desktop computer. In this task, instructors tell participants to remember the position of blue squares presented sequentially on the screen and touch a key when the current position matches that 1, 2, 3…n iterations back. Progression to blocks with increasing values of n is contingent on exceeding a threshold success rate. We measured n-back task performance as the highest n achieved.

4)  *Fluid intelligence* [@cattell1963] refers to the capacity to engage in abstract reasoning and problem solving in a way that is minimally dependent on prior experience. It complements "crystallized intelligence" (the ability to apply learned procedures and knowledge) as one of the two factors ($g_f$, $g_c$) comprising so-called "general intelligence" ($g$). Fluid intelligence is closely related to the executive control of attention and manipulation of the information held in working memory [@engle2018]. Researchers hypothesize that fluid intelligence supports technological innovation [@coolidge2005] and/or the intentional learning of new skills [@unsworth2005; @stout2015]. We measured fluid intelligence using the short version [@bilker2012] of the classic Raven Progressive Matrices task, which requires participants to complete increasingly difficult pattern matching questions.

5)  The *use of social information* for learning and decision-making varies across individuals and societies [@molleman2019]. Such variation is a key topic for understanding social learning and cultural evolutionary processes [@kendal2018; @heyes2018; @miu2020] and represents a potential confound for assessing experimental effects of different social learning conditions. We measured participants' tendency to rely on social information vs. their insights using the Berlin Estimate AdjuStment Task (BEAST) developed by Molleman et al. [-@molleman2019]. In this task, participants are present with large arrays of items on a screen and asked to estimate the number present. They are then provided with another person's estimate and allowed to give a second estimate. The participants' average adjustment between the first and second estimates measures their propensity to rely on social information.

## Stone Tool Making

After individual difference testing, participants engaged in a 2-hour stone tool making session, with a 15-minute break after 1 hour. During these breaks, participants were instructed not to seek additional training or information on stone tool-making (i.e., via the internet). For each group of participants, we randomly assigned subjects to one of two experimental conditions: no teaching or teaching. In both conditions, participants were first allowed to inspect and handle examples made by the expert model/instructor (**Figure** \@ref(fig:fig1)) of the kind of stone tools (flakes) they were asked to produce. They were told that their objective was to detach as many flakes as possible from the materials provided. These instructions meant that even the uninstructed condition included some minimal instruction (being told the objective). However, this was unavoidable without creating a more elaborate and naturalistic context where participants would develop their own technological goals. We expected that such a design would also increase behavioral variability, demanding correspondingly larger samples of participants to identify patterns and making direct comparisons with the instructed condition problematic.

```{r fig1, echo=FALSE, out.width = '100%', fig.align='center', fig.cap="Subjects examining demonstration flakes prior to the experiment. The demonstration lakes were made from the same basalt as used in the experiment with the same knapping technique."}
knitr::include_graphics(here::here("figure/Figure1.jpg"))
```

### Raw Materials

Each participant was provided with nine nodules for use over the 2-hour experiment. We made these nodules from larger chunks of a fine-grained basalt purchased from neolithics.com by fracturing them with a sledgehammer. This basalt has not been mechanically compared [e.g., rebound hardness, @braun2009] to East African basalts. Still, the basalt appears qualitatively like finer-grained examples at Gona (D. Stout, pers. obs.). Spalling produced irregular, angular chunks (**Figure** \@ref(fig:fig2)) for the experiment, weighing between 459g - 1876g (mean = 975g). All cores were weighed, measured (Length, Width, Thickness), and painted white so that we could discriminate new fracture surfaces from those created during production (see supplementary Table 1). Cores were sorted by shape and weight and then distributed evenly to each participant. As a result, there were no significant difference across participants in the mean weight (ANOVA, df = 22, F=0.3, p = 0.9; Levene test of homogeneity of variance = 1.04, df1=22, df2 = 184, p = 0.4) or shape (Length $\times$ Width/Thickness: ANOVA, df = 22, F=0.4, p = 0.9; Levene statistic = 0.6, df1=22, df2 = 184, p = 0.9) of cores provided. This was also true for cores provided to participants across the two experimental conditions (Instructed vs. Uninstructed mean weight = 1001g vs. 956g, t = 1.24, df = 205, p = 0.2, Levene's Test F = 0.6, p = 0.4; mean shape = 221.43 vs. 221.45, t = -0.003, df = 205, p = 0.9, Levene's Test F = 3.8, p= 0.05). Participants were, however, allowed to choose which cores to work on so that differences in the weight and shape of cores used across participants and conditions could still emerge because of selection bias.

```{r fig2, echo=FALSE, out.width = '100%', fig.align='center', fig.cap="3D scan of a basalt core prior to knapping. Cores were spray painted white to facilitate subsequent technological analyses of the flake scars and flaking intensity. Scale in the background is in 10mm blocks."}
knitr::include_graphics(here::here("figure/Figure2.jpg"))
```

Sixty pounds of 3-to-5 inch basalt "Mexican Beach Pebbles" were purchased from a landscaping supply company as hammerstones in the experiment. Of these, we selected 90 as suitable for use. These weighed between 213g-1360g (mean = 425) and varied in elongation (L/W = 1.01 to 2.65) and relative thickness (L$\times$W/T = 90.48 to 283.67). We placed roughly two-thirds of the stones in the middle of the knapping area (**Figure** \@ref(fig:fig3)) for participants to freely choose from during the experiment. Broken hammerstones were replaced from the reserve to maintain a consistent number and range of choices. We numbered each hammerstone and recorded each participants' choices along with the number of the core(s) they worked on with a particular hammerstone.

### Experimental Conditions

In both conditions, three researchers were present to record activities and collect materials. Participants were seated in a circle [@shea2015], and experiments were video recorded using two cameras. Participants were free to select hammerstones from the common pile and work on any of their nine assigned cores in any order they preferred (**Figure** \@ref(fig:fig3)). However, each core and all associated debitage were collected before participants were allowed to start working on a new core, so it was not possible to partially work and then return to a particular core later. The order of cores used and associated hammerstones were recorded for each participant during the experiment.

```{r fig3, echo=FALSE, out.width = '100%', fig.align='center', fig.cap="Subjects selecting hammerstones."}
knitr::include_graphics(here::here("figure/Figure3.jpg"))
```

In the uninstructed condition, a researcher (DS) sat with the participants and made stone tools but remained silent and made no effort to facilitate learning (e.g., through gesture, modified performance, facial expression, attention direction, or verbal instruction). Over the 2 hours, the researcher completely reduced four cores (one every \~30 minutes). We did not restrict participants from talking to each other, creating an unnatural and potentially stressful social context that might affect learning. Participants were asked to avoid any form of communication about the tool-making task specifically, and they complied with this request. Participants in this condition thus had the opportunity to observe tool-making (**Figure** \@ref(fig:fig4)) by an expert and/or by other learners, should they choose to do so, but received no intentional instruction.

```{r fig4, echo=FALSE, out.width = '100%', fig.align='center', fig.cap="Subjects observing expert knapper in the uninstructed condition."}
knitr::include_graphics(here::here("figure/Figure4.jpg"))
```

In the instructed condition, there were no restrictions on participant interaction, and the researcher engaged in direct active teaching [@kline2015] of tool-making techniques through verbal instruction, demonstration, gesture, and shaping of behavior. The instructor has a moderate level of experience teaching basic knapping skills to students in undergraduate archaeology classes and participants in previous knapping research [e.g., @stout2011]. The pedagogical strategy employed was based on the instructor's own learning experiences and theoretical interpretations [e.g., @pargeter2020] and focused on coaching participants in effective body postures, movement patterns, and grips and the assessment of viable core morphology.

## Lithic Analysis

**Table** \@ref(tab:tab1) lists details for the study’s various lithic attributes. We weighed and measured all the finished cores (L, W, T). Delta weight was calculated as (nodule start weight-core end weight)/nodule start weight. All detached pieces (DPs) were collected and weighed. We did not sort DPs into types (e.g., whole flakes, fragments), which would have greatly increased processing time. It is not clear that such distinctions add relevant information regarding utility/desirability beyond that supplied by metrics [@stout2019]. All DPs larger than 40mm in maximum dimension were photographed and measured. It is conventional in Early Stone Age lithic analysis to employ a 20 mm cut-off. We selected a higher threshold for both pragmatic (analysis time) and theoretical reasons. Flake use experiments have shown that flakes weighing less than  5–10 g or with a surface area below 7–10 $cm^2$ [@prasciunas2007] or with a maximum dimension \<50-60 mm [@key2014] become markedly inefficient for basic cutting tasks. Similarly, data from Oldowan replication experiments [@stout2019] show that the utility index (flake cutting edge/flake mass$^{1/3}$) \* (1 - exp[-0.31 \* (flake maximum dimension - 1.81)]) developed by Morgan et al. (2015) falls off rapidly below 40mm maximum dimension (Mean Utility \< 40mm = 0.508; \>=40mm = 0.946; t= 11.99, df = 707, p \< 0.000). By including weight in our cut-off criteria, we also avoid skewing the flake shape distribution by selectively retaining long, thin pieces (i.e., MD \> 40, weight \< 5g) while discarding rounder pieces of similar (or greater) weight and area. We recorded the absolute number of above-threshold DPs produced per core and calculated a mass-based proportion of larger DPs by dividing the combined weight of DPs \>40mm in maximum dimension and 5g in weight by the weight of all DPs. Higher values show cores from which proportionally more of the detached mass comprises large pieces.
```{r tab1, echo=FALSE, message=FALSE}
library(tidyverse)
df1 <- data.frame("Variable" = c("Delta mass", "Mass of large detached pieces / total debitage mass", "Length", "Width", "Thickness"), 
                 "Variable class" = c("Ratio", "Ratio", "Linear measurement", "Linear measurement", "Linear measurement"), 
                 "Definition" = c("Final core mass as a percentage of original nodule mass. Higher values show more completely flaked nodules.", "Combined mass of all detached pieces >40mm in maximum dimension and 5g in mass divided by the mass of all material detached from a specific core. Higher values show cores with more detached pieces per knapped material.", "Measurement along each core and flake's longest axis in mm.", "Measurement taken at 10% increments along a detached piece's longest axis starting at the detached piece's tip (in mm). Same measurement taken orthogonal to a core's length.", "Measurement taken at the maximum point in a detached piece's profile and orthogonal to width on the cores."),
                "Recording method" = c("Digital scale", "Calipers, counts, digital scale", "2D photogrammetry", "2D photogrammetry", "Calipers") )

kableExtra::kbl(df1, booktabs = T, caption = "Overview of lithic variables and recording methods used in the study.") %>%
  kableExtra::kable_styling(latex_options = c("striped", "HOLD_position"), font_size = 7) %>%
  kableExtra::column_spec(1, width = "10em") %>%
  kableExtra::column_spec(3, width = "20em")%>%
  kableExtra::column_spec(4, width = "10em")

```

We defined DP length as the longest axis and width as the maximum dimension orthogonal to length for measurement (see supplementary Table 2). Our thickness measurement follows the maximum dimension orthogonal to the plane formed by L and W, which we measured using calipers. L, W, and plan-view area measurements were taken from photographs captured using a Canon Rebel T3i fitted with a 60 mm macro lens and attached to a photographic stand with adjustable upper and lower light fittings. The camera was positioned directly above the flakes and kept at a constant height. We arranged DPs irrespective of any technological features so that the longest axis was vertical and placed the broader end toward the bottom of the photograph.

* Photographs were post-processed using Equalight software to adjust for lens and lighting falloff resulting from bending light through a lens and its aperture, affecting measurements taken from photos. We shot each image with a scale then used this scale to rectify the photograph's pixel scale to a real-world measurement scale in Adobe Photoshop. Images were converted to binary black and white format, and silhouettes of the tools were extracted in Adobe Photoshop. We then used a custom ImageJ [@rueden2017] script [@pargeter2019]to measure DP length and take nine width measurements at 10% increments of length starting at the tip of each DP (i.e., 10% width = tip, 90% width = base). We used the built-in ImageJ tool to measure DP area.

## Statistical Analyses

We adopted an information-theoretic approach to evaluate the association between psychometric, motor-skill, training measures, and technological outcomes [@burnham2002]. Information-theoretic techniques provide methods for model selection using all possible combinations of variables while avoiding problems associated with significance-threshold stepwise selection. We used the small sample corrected Akaike information criterion (AICc) to rate each possible combination of predictors on the balance between the goodness of fit (likelihood of the data given the model) and parsimony (number of parameters). The AICc consists of the log-likelihood (i.e., how well does the model fit the data?) and a penalty term for the number of parameters that must be estimated in the model (i.e., how parsimonious is the model?), with a correction for small sample sizes (AICc converges to the standard AIC at large samples). A lower AICc indicates a more generalizable model, and we use it to compare and rank various possible models. Each analysis begins with a complete model that includes all predictors of interest. All possible combinations of predictors are then fitted, and the resulting models are ranked and weighted based on their AICc. We chose the "best" model because it had the lowest AICc score.

Continuous predictors were centered such that zero represents the sample average, and units are standard deviations. The full model was fitted with the *lm* function in R 4.0.3, and the *glmulti* package (Bartoń 2015) was used for multi-modal selection and model comparison.

# Results

Following a recent protocol to enhance the reproducibility and data transparency of archaeological research [@marwick2017], detailed results of all analyses and assessments of the data structure are available in our paper's supplementary materials and through Github (<https://github.com/Raylc/PaST-pilot>).  Here we limit discussion to the major ﬁndings regarding flaking performance and individual differences. We were particularly interested in 1) group-level effects of experimental conditions, 2) individual differences in aptitude and learning, and 3) potential interactions between learning conditions and individual differences. To address these questions, we employed data reduction (Principal Component Analysis) to derive two summary metrics of flaking performance, compared these factors across the two experimental conditions, and built multivariate models examining the relations between our various psychometric measures, subject's motor skill scores, and our two lithic performance factors.

## Principal Component analyses

The following two sections outline factor analyses that summarize our main study metrics tracking individual variation in DP sizes and shapes and lithic performance measures.

### Detached Piece size and shape

To better understand the relationship between DP shape and training/individual variation, we entered our nine linear plan measurements and maximum length and thickness into a principal component analysis (PCA) from which we extracted summary coordinates. Bartlett's Test of Sphericity was significant ($\chi ^2(10)$  = 4480, p \< .01), indicating that the set of variables are adequately related for factor analysis.

The analysis yielded three factors explaining a total of 90% of the variance for the entire 11 measurement variable set (**Table** \@ref(tab:tab2)). Factor 1 tracks size with higher scores indicating larger pieces since all 11 measures positively load this factor. Factor 2's loadings track increasing thickness and length but not width. As factor 2 scores increase, pieces thus get thicker, longer, and relatively narrower, resembling irregular splinters. Factor 3 tracks increasing proximal breadth and decreasing thickness. As factor 3 scores go up, pieces get thinner overall and broader at the base. Factor 3, therefore, tracks pieces with a more typical flat, flake-like morphology having a thin cross-section and tapering shape. We used these three shape coordinates to approximate DP size and shape in the project's flaking performance factor analysis.
```{r tab2, echo=FALSE, message=FALSE}
library(tidyverse)
df2<-read.csv(here::here("data/tables/table2.csv"), fileEncoding = 'UTF-8-BOM')
kableExtra::kbl(df2, booktabs = T, caption = "Lithic size/shape PCA factor loadings.") %>%
  kableExtra::kable_styling(latex_options = c("striped", "HOLD_position"), font_size = 7)%>%
kableExtra::row_spec(0, bold = T)%>%
kableExtra::row_spec(1, bold = T)
```

### Lithic flaking performance measures

We conducted a second principal component analysis better to understand the relationship between our various lithic performance measurements and reduce data dimensionality. This analysis examined the study's six lithic performance measures (count of large pieces [\>40mm and 5g], mass of large pieces relative to total detached mass, core delta mass, and the three flake shape factors) (see supplementary Table 3 for details). These measures were summarized for each core, and individual factor scores were calculated from these core-specific measures. Bartlett's Test of Sphericity was significant ($\chi ^2(6)$  =3185, p \< .01) indicating that the set of variables are at least adequately related for factor analysis.

The analysis yielded two factors explaining 56% of the variance for the entire set of variables (**Table** \@ref(tab:tab3)). Factor 1 (hereafter "Quantity") explains 28.7% of the variance and tracks flaking quantity due to high positive loadings on large DP count and mass ratio and core delta mass. Performance factor 2 (hereafter "Quality") covers 27% of the sample variance and measures flaking quality as reflected in high positive loadings on Shape Factors 1 (size) and 3 (thin, "flake-like" shape) and a negative loading on Shape Factor 2 ("splinter-like" thickness and elongation). High scores on Quality thus reflect the making of larger, relatively thinner, and more typically flake-shaped vs. splinter-shaped DPs.
```{r tab3, echo=FALSE, message=FALSE}
library(tidyverse)
df3<-read.csv(here::here("data/tables/table3.csv"), fileEncoding = 'UTF-8-BOM')
kableExtra::kbl(df3, booktabs = T, caption = "Overview of principal components results on the six flake performance measures. DP = detached piece.") %>%
  kableExtra::kable_styling(latex_options = c("striped", "HOLD_position"), font_size = 7)%>%
kableExtra::row_spec(0, bold = T)%>%
kableExtra::row_spec(1, bold = T)
```

These two factors address flaking performance at the level of individual cores. However, we were also interested in each participant's overall productivity/work rate over the entire two hours. For example, looking at a knappers average Quality and Quantity factor scores would not differentiate between a participant who spent the whole time exhaustively reducing one core vs. another participant who did the same to all nine of their allotted cores in the same time. We calculated a simple Total Productivity metric as the sum of all mass a participant removed from cores during the experiment to capture this aspect of variation.

## Relationships between Performance Measures

This approach also allowed us to compare the relationship between Total Productivity, Quantity, and Quality across our two experimental groups (**Figures** \@ref(fig:fig5) and \@ref(fig:fig6)). As might be expected, we found that per-core Quantity and Total Productivity are positively correlated in both groups (**Figure** \@ref(fig:fig5)a), although this relationship is twice as strong in the instructed  (F[1, 9] = 33, p \< 0.01, Adj. $R^2$ = 0.8) compared to uninstructed  (F[1, 8] = 8, p = 0.02, Adj. $R^2$ = 0.4) group. Interestingly, we also found evidence of a negative correlation between Total Productivity and Quality in the uninstructed group (F[1, 8] = 28, p = \<0.01, Adj. $R^2$ = 0.7), but no relation in the instructed group (**Figure** \@ref(fig:fig5)b). A qualitatively similar trend concerning Quality vs. Quantity (**Figure** \@ref(fig:fig5)c) did not achieve significance (F[1, 17] = 0.6, p = 0.2).

```{r fig5, echo=FALSE, out.width = '100%', fig.align='center', fig.cap="Relationships between flaking quantity, quality, productivity, and the two training conditions. Each dot represents a participant, colors represent training conditions."}
knitr::include_graphics(here::here("figure/Figure5.jpg"))
```

Thus, it appears that instructed participants achieved higher Total Productivity by increasing average flaking Quantity across cores and without sacrificing Quality. In contrast, uninstructed participants found other ways to vary Total Productivity (e.g., number of cores knapped rather than Quantity per core, see **Table** \@ref(tab:tab5), **Figure** \@ref(fig:fig8) & \@ref(fig:fig9)) and generally increased productivity at the expense of Quality. Experimental artifacts illustrating these trade-offs are presented in **Figure** \@ref(fig:fig6).

```{r fig6, echo=FALSE, out.width = '100%', fig.align='center', fig.cap="Comparison of untrained and trained detached pieces illustrating the ability of Trained participants to maximize Quality and Productivity at the same time."}
knitr::include_graphics(here::here("figure/Figure6.jpg"))
```

## Do trained, untrained, and expert knappers perform differently?

Here we compare our flaking outcomes (DP size/shape and flaking performance factors) between the instructed and uninstructed groups. We include the expert demonstrator/instructor as a performance benchmark.

**Table** \@ref(tab:tab4) summarizes the results of ANOVA tests group-level difference in central tendency on various performance measures. We found no significant differences between the instructed and uninstructed groups on our flaking Quantity and Quality factors. In contrast, three-way flake size and shape comparisons between our expert knapper and the two novice groups showed that the expert knapper made significantly more large flakes (effect size = 0.14). The expert also had a significantly higher core delta mass signal than either of the novice groups (effect size = 0.26) and left significantly smaller finished cores (effect size = 0.27) (**Figure** \@ref(fig:fig7)). All three of these results show either medium or large effect sizes. The instructed group's data distributions tended towards the expert sample in all three comparisons, although they did not differ significantly from the uninstructed group (**Figure** \@ref(fig:fig7)). We also observed a significant difference in shape factor 2 (splinters) driven by the expert's lower values but with a very small (\< 0.01) effect size. These results show that mean core reduction intensity and large flake production rates distinguish expert and novice performance. In contrast, novices in experimental groups produced similar mean detached piece sizes and shapes as the expert trainer. 
```{r tab4, echo=FALSE, message=FALSE}
library(tidyverse)
df4<-read.csv(here::here("data/tables/table4.csv"), fileEncoding = 'UTF-8-BOM')
kableExtra::kbl(df4, booktabs = T, caption = "Summary group-wise comparison statistics contrasting flake performance metrics by experiment condition and with the expert instructor. All comparisons are three-way except for the two skill PC factors and the total mass removed, which are two-way comparisons between the subject population as these factors were not applied to the expert data sample. Eta2 = ANOVA effect size (>0.1 = medium effect size, >0.2 = large effect size). * = non-parametric permuted p-values to account for unequal sample variances. SS = sum of squared differences, df = degrees of freedom, MS = mean squared difference.") %>%
  kableExtra::kable_styling(latex_options = c("striped", "HOLD_position"), font_size = 7)%>%
kableExtra::row_spec(0, bold = T)
```

```{r fig7, echo=FALSE, out.width = '100%', fig.align='center', fig.cap="Results showing significant differences between the instructor (expert) and novice flaking performance."}
knitr::include_graphics(here::here("figure/Figure7.jpg"))
```

```{r fig8, echo=FALSE, out.width = '100%', fig.align='center', fig.cap="Comparison of uninstructed, instructed, and expert cores."}
knitr::include_graphics(here::here("figure/Figure8.jpg"))
```

While we did not find significant differences in central tendency between our two experimental groups, results (**Figure** \@ref(fig:fig7)) did indicate lower variance in the instructed group. To test whether training reduced variability in performance outcomes between subjects, we compared variance metrics between the instructed and uninstructed individuals using the F-test on either core-averaged or flake-specific variances. **Table** \@ref(tab:tab5) and **Figure** \@ref(fig:fig9) present the results from these comparisons showing significant variance differences predominantly in flaking Quality, number of large DPs, core delta mass, and the total amount of flaked mass). In most instances, variance in the uninstructed group exceeds that of instructed individuals by 1.5 to 4.7 times. The most salient effect of instruction was thus not to shift mean performance but to reduce variability by eliminating the skew (generally toward poorer outcomes) seen in the uninstructed group (**Figure** \@ref(fig:fig9)), rather than to shift the mean.
```{r tab5, echo=FALSE, message=FALSE}
library(tidyverse)
df5<-read.csv(here::here("data/tables/table5.csv"), fileEncoding = 'UTF-8-BOM')
colnames(df5)[colnames(df5) %in% c("X95..CI")] <- c("95% CI")
kableExtra::kbl(df5, booktabs = T, caption = "F-test results comparing variances between trained and untrained subjects across 
our various flaking performance metrics.") %>%
  kableExtra::kable_styling(latex_options = c("striped", "HOLD_position"), font_size = 7)%>%
kableExtra::row_spec(0, bold = T)
```

```{r fig9, echo=FALSE, out.width = '100%', fig.align='center', fig.cap="Variance comparisons between instructed and uninstructed individuals across various flaking performance metrics. Note the lower variance of the trained group in all the plots."}
knitr::include_graphics(here::here("figure/Figure9.jpg"))
```

## Does performance change over time?

In addition to comparing overall performance during the two-hour experiment, we also wanted to determine if groups or individuals differed in learning (i.e., performance change) over the period.  We calculated the learning stage as the ordinal number of each core out of the total number knapped by each subject (i.e., core 2 of 4 or 4 of 8, both equal 50% complete) for these analyses. These relative core use-order percentages were then binned into 20 percent brackets for core-order and group-level comparisons. We tracked flaking outcomes using the two performance factors (Quality and Quantity). We added the nodule starting mass to track whether training/practice times impacted raw material selection.

**Table** \@ref(tab:tab6) shows no significant training effects across the two performance measures either as grouped data or between individuals (**Figure** \@ref(fig:fig10) and **Figure** \@ref(fig:fig11)). This result demonstrated that flaking outcomes did not change dramatically across the study interval. An inspection of individual learning curves confirms this lack of significant learning effects (**Figure** \@ref(fig:fig11)). The main training effect was core starting mass (with a substantial effect size = 0.25). On average, core starting masses start low and increase, showing that participants selected smaller nodules first. As participants depleted the smaller nodules in their allotment, they knapped larger, less preferred nodules. This preference for smaller cores is somewhat less pronounced in the uninstructed group, as indicated by a small main effect of learning condition and generally higher starting nodule masses for the uninstructed group (**Figure** \@ref(fig:fig11)). 
```{r tab6, echo=FALSE, message=FALSE}
library(tidyverse)
df6<-read.csv(here::here("data/tables/table6.csv"), fileEncoding = 'UTF-8-BOM')
kableExtra::kbl(df6, booktabs = T, caption = "Summary group-wise comparison statistics contrasting flake performance metrics by experiment condition and relative core order. All comparisons are three-way. Eta2 = ANOVA effect size (>0.1 = medium effect size, >0.2 = large effect size). SS = sum of squared differences, df = degrees of freedom, MS = mean squared difference.") %>%
  kableExtra::kable_styling(latex_options = c("striped", "HOLD_position"), font_size = 7)%>%
kableExtra::row_spec(0, bold = T)
```

```{r fig10, echo=FALSE, out.width = '100%', fig.align='center', fig.cap="Comparisons of nodule starting mass across the study period by training condition. Results show a significant relationship between nodule starting mass and length of time in the study."}
knitr::include_graphics(here::here("figure/Figure10.jpg"))
```

```{r fig11, echo=FALSE, out.width = '100%', fig.align='center', fig.cap="Individual learning curves derived from each subject’s relative core use order and the two flake performance factors."}
knitr::include_graphics(here::here("figure/Figure11.jpg"))
```

## Do individual differences in motor skill and psychometric measures predict flaking performance?

One of the experiment's primary goals was to test if individual perceptual-motor and cognitive variation measures predict success in stone flaking across different training conditions. We built three multivariate models to address this goal examining the relations between training conditions, individual difference measures, and our three lithic performance measures (overall productivity and average per-core Quantity and Quality). These models enabled us to determine which psychometric and motor skill factors are better predictors of a participant's flaking performance in the study.

We considered all possible interactions between five individual difference measures, core size, training condition, and the three performance measures, with each subject providing one data point. Each model's continuous predictors (highest n-back level, Raven's Progressive Matrix score, BEAST score, starting nodule mass, Fitts score, and grip strength) were centered such that zero represents the sample average, and units are standard deviations. Our two motor skill and strength measures (grip strength and Fitt’s performance scores) are also strongly correlated (F [1,19] = 15, p \< 0.01, $R^2$ = 0.41). However, these two measures track complementary components of athleticism (strength vs. speed/accuracy trade-offs), so we decided to include both in the model selection process.

All three models follow the same complete model statement as follows:

*Flaking performance variable $\sim$ Training condition + Highest n-back level + Raven's Progressive Matrix score + BEAST score + Fitt's score + Grip strength*

For our two per-core performance factors (Quantity and Quality), it is also relevant to consider how individual core features may have affected performance. We found no evidence of individual or group level practice effects over the two hours, so we did not include core order in the models. We did, however, find that subjects selected progressively larger nodules throughout the experiment. It is thus essential to understand whether nodule variability had any impact on our flaking results. Because starting nodule size (mass) and shape were strongly correlated (F [1,157] = 186, p \< 0.01, $R^2$ = 0.54), we included nodule mass as a covariate to control for any variance in flaking performance that nodule differences may drive.

### Model 1: Individual differences and overall productivity

Our first model examined variance in overall flaking productivity measured by each subject's combined flaked mass (nodule starting mass-core final mass). This productivity metric provides a basic measure of variation in individuals' success detaching pieces and reducing cores from a standardized (see Methods) raw material supply. From the same candidate pool size of 55893 possible multivariate models, the best performing model returned an AICc value of -18 (Average AIC = -13). This model comprised the following statement with two main and three interaction effects:

*Total flaked mass* $\sim$ Training condition + Grip strength + RPM$\times$Highest n-back level + Fitts score$\times$BEAST score + Grip strength$\times$RPM

This model explains a statistically significant and substantial proportion of variance in flaking productivity ($R^2$ = 0.84, F (6, 14) = 12.7, p \< 0.01, adj. $R^2$ = 0.77). A model residuals normality test shows no significant differences with the normal distribution (p = 0.72) indicating that this relationship is linear. A Breusch-Pagan test showed no evidence for heteroskedasticity (BP = 2, df = 6, p = 0.8).
```{r tab7, echo=FALSE, message=FALSE}
library(tidyverse)
df7<-read.csv(here::here("data/tables/table7.csv"), fileEncoding = 'UTF-8-BOM')
colnames(df7)[colnames(df7) %in% c("X95..CI", "t..12.")] <- c("95% CI", "t(12)")
kableExtra::kbl(df7, booktabs = T, caption = "Summary and estimates for the overall flaking productivity model.") %>%
  kableExtra::kable_styling(latex_options = c("striped", "HOLD_position"), font_size = 7)%>%
kableExtra::row_spec(0, bold = T)
```

**Table** \@ref(tab:tab7) presents this model's coefficients and summary outputs, wherein baseline refers to the uninstructed condition with all continuous predictors at the sample average. The parameter estimates for the continuous predictors reflect the expected change in utility for one standard deviation change in the predictor variable. We found significant (p\< 0.05) and substantial (Standardized Estimate \>= 0.50, i.e. a 50% change in variable) main effects of Grip Strength, Visuospatial nBack, and BEAST. The main effect of Grip Strength (**Figure** \@ref(fig:fig12)), irrespective of learning condition, indicates the basic importance of strength in generating higher production rates among naive knappers, at least when we do not consider efficiency and quality.
```{r fig12, echo=FALSE, out.width = '100%', fig.align='center', fig.cap="Significant main effect of grip strength on overall flaking productivity."}
knitr::include_graphics(here::here("figure/Figure12.jpg"))
```

Effects of visuospatial working memory capacity and social information use are more complicated, as indicated by strong interactions with learning condition (**Figure** \@ref(fig:fig13)). In each case, higher scores were associated with better performance in the uninstructed group but *worse* performance in the instructed group. Positive effects in the uninstructed group were as expected, given the hypothesized importance of spatial cognition [@coolidge2005] and social learning [@morgan2015] in acquiring knapping skills. Negative effects in the instructed group are unexpected but presumably reflect differences in learning strategies adopted under instruction.

```{r fig13, echo=FALSE, out.width = '100%', fig.align='center', fig.cap="Two significant interaction effects of training condition, social information, and visuo-spatial working memory on overall flaking productivity."}
knitr::include_graphics(here::here("figure/Figure13.jpg"))
```

### Model 2: Individual differences and quality flaking

The second full model examined the variance in average flaking Quantity per core. It thus complements our first model assessing overall productivity by testing for differences in reduction intensity at the level of individual cores. From a candidate pool of 55893 possible multivariate models, the best performing model returned an AICc value of 32 (Average AIC = 44). This model comprised the following statement with three main and four interaction effects:

*Quantity $\sim$ Highest n-back level + BEAST score + Fitt's score + Grip strength + Training condition$\times$Highest n-back level + Training condition$\times$BEAST score + Training condition$\times$Grip strength + Nodule mass (as control)*

This model explains a statistically significant and substantial proportion of variance in quantity flaking ($R^2$ = 0.7, F (8, 12) = 3.6, p = 0.02, adj. $R^2$ = 0.5). A model residuals normality test shows no significant differences with the normal distribution (p = 0.38) indicating that this relationship (as required) is linear. A Breusch-Pagan test showed no evidence for heteroskedasticity (whether variance for all observations in our data set are the same) (BP = 4.4, df = 8, p = 0.8).

**Table** \@ref(tab:tab8) presents this model’s coefficients and summary outputs, following the same format as **Table** \@ref(tab:tab6) and **Table** \@ref(tab:tab7).
```{r tab8, echo=FALSE, message=FALSE}
library(tidyverse)
df8<-read.csv(here::here("data/tables/table8.csv"), fileEncoding = 'UTF-8-BOM')
colnames(df8)[colnames(df8) %in% c("X95..CI", "t..12.")] <- c("95% CI", "t(12)")
kableExtra::kbl(df8, booktabs = T, caption = "Summary and estimates for Quantity flaking model.") %>%
  kableExtra::kable_styling(latex_options = c("striped", "HOLD_position"), font_size = 7)%>%
kableExtra::row_spec(0, bold = T)
```

The Quantity model roughly paralleled results for Total Production, yielding substantial and significant interactions between training condition, n-back level, BEAST scores, and grip strength. As with Total Production, higher visuospatial n-back levels and BEAST scores were associated with lower Quantity scores in the instructed group but higher or unchanged DP Quantity in the uninstructed group (**Figure** \@ref(fig:fig14)).

```{r fig14, echo=FALSE, out.width = '100%', fig.align='center', fig.cap="Two significant interaction effects of training condition, visuo-spatial working memory, and social information, on overall flaking quality."}
knitr::include_graphics(here::here("figure/Figure14.jpg"))
```

Unlike Total Productivity, an interaction with learning condition mediated the effect of Grip Strength on per-core Quantity (**Figure** \@ref(fig:fig15)). Only instructed individuals translated Grip Strength into a more intense reduction of individual cores, including delta mass and the number and proportion of larger pieces. Thus, high Grip Strength enabled individuals in both groups to produce more debitage.

```{r fig15, echo=FALSE, out.width = '100%', fig.align='center', fig.cap="Three significant interaction effects of training condition, visuo-spatial working memory and social information, and training and grip strength use on quantity flaking."}
knitr::include_graphics(here::here("figure/Figure15.jpg"))
```

### Model 3: Individual differences and quality flaking

Our third model examining variance in Quality follows the same complete model statement we used for Quantity. From the same candidate pool size of 55893 possible multivariate models, the best performing model returned an AICc value of 32 (Average AIC = 39). This model comprised the following statement with three main and four interaction effects:

*Quality flaking $\sim$ Highest n-back level + Fitt's score + Grip strength + Fitt's score$\times$BEAST score + Grip strength$\times$BEAST score + Grip strength$\times$Fitt's score + Training condition$\times$Grip strength + Nodule mass (as control)*

This model explains a statistically significant and substantial proportion of variance in Quality ($R^2$ = 0.75, F (8, 12) = 4.6, p \< 0.01, adj. $R^2$ = 0.6) in the absence of any main training effects. A model residuals normality test shows no significant differences with the normal distribution (p = 0.41) indicating that this relationship is linear. A Breusch-Pagan test showed no evidence for heteroskedasticity (BP = 7, df = 8, p = 0.5).
```{r tab9, echo=FALSE, message=FALSE}
library(tidyverse)
df9<-read.csv(here::here("data/tables/table9.csv"), fileEncoding = 'UTF-8-BOM')
colnames(df9)[colnames(df9) %in% c("X95..CI", "t..12.")] <- c("95% CI", "t(12)")
kableExtra::kbl(df9, booktabs = T, caption = "Summary and estimates for the Quality flaking model.") %>%
  kableExtra::kable_styling(latex_options = c("striped", "HOLD_position"), font_size = 7)%>%
kableExtra::row_spec(0, bold = T)
```

**Table** \@ref(tab:tab9) presents this model’s coefficients and summary outputs following the same data format as **Table** \@ref(tab:tab7) & **Table** \@ref(tab:tab8). The results show no effects that were both significant and substantial (Estimate >= 0.5).

The Quality model did produce two statistically (p\< 0.05) significant interaction effects (RPM \* BEAST & Fitts \* n-back). However, these interactions had relatively small effects on Quality (\<0.5), and we believe that interpreting these results from our small, exploratory study would be inappropriate. **Figure** \@ref(fig:fig16) shows the uneven distribution of data points for these interactions, which suggests vulnerability to leveraging effects of a small number of extreme value combinations.

```{r fig16, echo=FALSE, out.width = '100%', fig.align='center', fig.cap="Significant interactions with low estimated effect sizes from the quality flaking model."}
knitr::include_graphics(here::here("figure/Figure16.jpg"))
```

Without a larger sample, it is not possible to determine if these are anomalous outliers or simply represent a poorly sampled part of the broader population.

## Behavioral observations

We designed this exploratory study primarily to trial experimental design elements such as training time, conditions, and raw materials and to collect preliminary data on individual differences and training effects on knapping outcomes. We thus focused on collecting quantitative psychometric and lithic data. However, we also considered that quantifying participant knapping behaviors and products could be important for future studies. To support methods development in this regard, we made ad hoc notes on observed behaviors during the experiments and video-recorded all experiments to enable later, more systematic analyses yet to be completed. However, even casual behavior observation was sufficient to reveal an unexpected effect. Whereas all instructed participants copied the general posture and technique of the expert (freehand knapping seated in a chair), fully half (6) of the uninstructed participants experimented with or even knapped all their cores using the floor as support (**Figure** \@ref(fig:fig17)). Three of these participants were in the same session, the only group composed of just three individuals. In this group, knapping on the ground appears to have been transmitted from one participant to the other two based on appearance order and the point of gaze of participants.

```{r fig17, echo=FALSE, out.width = '100%', fig.align='center', fig.cap="Novices in the uninstructed condition developing and transmitting a form of bipolar technology involving core reduction on the floor."}
knitr::include_graphics(here::here("figure/Figure17.jpg"))
```

# Discussion

The most salient finding of our exploratory study is that the presence/absence of teaching clearly impacted knapping performance but did so in nuanced and individually variable ways that previous research has not explored. We did not observe any significant differences in mean performance between our experimental conditions. Some non-significant tendencies toward enhanced instructed group performance suggest that a larger participant sample might detect significant effects, but also that the size of any such effects would likely remain small. This result could reasonably lead one to conclude that teaching does not substantially facilitate early-stage knapping skill acquisition [@putt2014; @ohnuma1997]. Looking closer, however, it is evident that teaching has several significant effects.

## Variance Reduction

In our experiment, the strongest effects of teaching were to reduce variance (**Figure** \@ref(fig:fig9), **Table** \@ref(tab:tab5)) rather than shift mean values. Instruction acted as a "safety net" that homogenized performance by reducing the frequency of inferior outcomes (i.e., learning failures). This finding provides additional support for the hypothesis that teaching would have increased the reliability of Oldowan skill reproduction [@morgan2015] while simultaneously corroborating the view that naïve knappers can achieve basic flaking competence in its absence [@tennie2017]. Our results thus do not imply that novices required teaching or that teaching was even present during Oldowan times (@gärdenfors2017). Instead, the results serve to reinforce the plausibility of co-evolutionary scenarios positing the cost/reliability of technological skill acquisition as a selection pressure favoring the evolution of teaching and language [@stout2010; @morgan2015; @stout2017].

## Knapping Behaviors

The current study complements the transmission chain design of Morgan et al. [-@morgan2015] (2015) by finding similar effects in more naturalistic learning contexts. Our design further allowed us to examine individual variation to understand better how teaching produces its effects. Whereas transmission chains are optimized to investigate iterative learning effects [see @caldwell2020], they necessarily involve each participant's different instructor/model. We sacrificed this iterative component to consider how the presence/absence of teaching affected the behavior of individuals under otherwise standardized learning conditions.

We found that a critical impact of teaching was to alter basic flake production strategies, as reflected in the relationship between Total Productivity and detached piece Quality (**Figure** \@ref(fig:fig5)b). Whereas uninstructed participants achieved greater Productivity at the expense of Quality and vice versa, these dimensions were unrelated across instructed participants. Thus, even though uninstructed participants gained the highest values on each metric, only instructed individuals managed to maximize both simultaneously. Indeed, a core function of teaching is to reduce the search space that learners must explore and increase the likelihood of discovering globally instead of locally optimum solutions [cf. @hinton1996; @stout2013]. In our study, uninstructed individuals explored a greater range of basic behavioral variations not seen in the instructed group, including knapping on the floor, concentrating on working just a few cores (2-4, **Figure** \@ref(fig:fig9)c) over the practice period. Uninstructed individuals also showed less constrained nodule size preferences (**Figure** \@ref(fig:fig8)). Notably, this variation occurred even in the presence of an observable expert example, suggesting it may be interesting for future experiments to address the impact of social context, expectations, and relationships on observational learning strategies [@kendal2018].

We also found a strong positive effect of Grip Strength on Total Productivity independent of instruction (**Figure** \@ref(fig:fig12)). While it is tempting to interpret this with respect to the demands for hand strength specifically, it is important to remember that grip strength correlates with total muscle strength [@wind2010] and overall fitness [@sasaki2007]. Thus, it is best taken to indicate some importance of fitness generally in increasing the rate and intensity of core reduction by naïve knappers, potentially affecting the rate of work and the kinetic energy of the swing and the handling of core and hammerstone. It thus provides further support for hypotheses positing stone tool-making as a selection pressure on the functional anatomy of hand, arm, and shoulder [e.g., @williams-hatala2018], but initially appears orthogonal to variations in learning condition and knapping behaviors in our study. However, we also found that the effect of Grip Strength on per-core knapping Quantity is dependent on teaching (**Figure** \@ref(fig:fig15)). The absence of this effect in the uninstructed group reflects the weaker association between Total Productivity and per-core Quality across these participants (**Figure** \@ref(fig:fig5)a). Grip Strength increased uninstructed Productivity specifically by allowing them to knap more cores rather than to reduce individual cores more heavily. In keeping with this, uninstructed Grip Strength positively correlates with Total Cores knapped  ($R^2$ = 0.54, p = 0.01). Conversely, strength allowed instructed participants to increase their average Quantity per core without affecting the total number of cores knapped ($R^2$ = 0.18, p = 0.165). Thus, strength appears to have affected core reduction rate and intensity in different ways, depending on teaching. This difference is likely related to the homogenizing effect of teaching on knapping rate (all instructed participants knapped six or more cores) and methods. Subjectively, uninstructed participants’ knapping behaviors often appeared more physically demanding (e.g., a greater number of non-productive blows, rapid and unregulated battering). This result would imply different demands on strength and aerobic fitness [@williams-hatala2021; @mateos2019]. However, future research will need to test this observation systematically.

In this respect, it is also important to note that we do not know how well the knapping objectives and strategies communicated by the expert in our experiment correspond to actual Oldowan goals and behaviors. The instructor has successfully replicated assemblage-level patterning at Gona [@stout2019]. Still, Oldowan behavior is variable across space and time [e.g., @braun2019], and alternative knapping methods might maximize different values (productivity, quality, effort), especially in novices [@putt2015]. Nevertheless, the effect of instruction to constrain behavioral exploration and homogenize outcomes is clear. We expect this effect to generalize to teaching alternative knapping goals and behaviors, although future work should test this prediction.

## Learning Strategies

One primary goal of this experiment was to test the viability of a moderate, two-hour learning period for skill acquisition studies. Unfortunately, we found that this duration was insufficient to capture learning effects for Oldowan-like flake production. We cannot attribute the lack of performance change over the period (**Figure** \@ref(fig:fig11)) to a ceiling effect (i.e., rapid task mastery at the outset of the practice period). Participants remained well below expert levels and continued to display the high within-individual variability typical of naïve/novice knapping [@eren2011; @pargeter2019]. This negative result was unexpected but is broadly consistent with evidence that Early Stone Age flaking, while conceptually simple, requires substantial practice for perceptual-motor skill development [@nonaka2010; @stout2015; @pargeter2020]. Future investigations of learning variation across individuals and/or experimental conditions may thus need to incorporate more extended practice periods to capture skill acquisition processes. In theory, future experiments might use shorter knapping trials to assess the variation in initial performance across individuals and under different conditions captured in our study. However, the presence of substantial core-to-core variation within individuals cautions against overly brief experiments that might not provide a representative sample. Greater durations also allow for the expression of different learning strategies over time, even in the absence of directional performance change.

At a basic level, learners of any new task must balance investment in task exploration vs. exploitation of knowledge and skills already in hand [@sutton2018]. Premature exploitation risks settling for a sub-optimal local solution, whereas continued exploration sacrifices more immediate payoffs. Managing this trade-off is especially challenging for complex, real-world tasks like stone knapping and is thought to depend on the interplay of uncertainty and reward expectation [@wilson2021]. Teaching and other social learning resources have the potential to provide low-cost information about task structure and payoffs [@rendell2010; @kendal2018], which, if adopted, would be expected to affect exploration/exploitation decisions. Such adoption is itself known to be influenced by individual cognitive differences, for example, if higher fluid intelligence allows observers to better understand observed tasks [@vostroknutov2018] or if individuals vary in their tendency to use and value social information [@toelch2014; @molleman2019].

Our study did not observe any effect of fluid intelligence (RPM) on knapping outcomes. Still, we found strong learning condition interactions with participant visuospatial working memory and social information use tendency (**Figure** \@ref(fig:fig13)). As expected, uninstructed individuals with higher scores on these dimensions displayed higher Total Productivity and average per-core flaking Quantity (although the effect on n-Back on Quantity did not achieve significance). We attribute these effects to an increased ability to hold relevant morphological/spatial information in mind and a tendency to benefit from observing successful strategies of others, including the expert model. In contrast, instructed individuals with higher scores tended to have lower Productivity and Quantity. We interpret this unexpected effect as an increased tendency to privilege exploratory learning behavior overexploitation. We suggest that instructed participants might knap more slowly and less productively if higher working memory capacity inclined them to experiment more with morphological/spatial variables highlighted by the instructor. Instructed subjects’ predisposition to use social information might have caused them to invest greater time and effort attending to and trying out observed actions and/or instructions. These suggestions remain to be tested by further work.

Putt et al. [-@putt2014] report a similar negative effect of instruction on knapping outcomes during early-stage learning. They interpret this result to reflect learners experimenting with advanced techniques before they have the perceptual-motor skill to execute them [@whiten2015; @stout2015]. Unfortunately, the training period in our current experiment was insufficient to capture learning effects, so we have no evidence of the impact of these individual differences and putative exploration/exploitation trade-offs to the ultimate achievement of expertise. Future research might explore such effects with more detailed behavioral data instead of purely lithic data and with more extended learning periods.

## Limitations and Prospects

Our exploratory study produced several robust results concerning the effects of instruction and individual differences on lithic products. However, we had a small sample size, short training duration, and a lack of detailed quantification of observed behaviors. We can hopefully address these limitations in future studies building on the methods and evidence presented here. For example, it is notable that our study failed to document any reliable effects on knapping Quality. This lack of a result might reflect an actual lack of such effects. Still, it may also indicate a need for more sensitive measures and/or increased sample size and training duration to identify subtle or delayed effects. One aspect of our attempt to balance pragmatic costs and benefits in our study was to test the efficacy of relatively limited lithic analysis. More detailed ongoing investigations of core morphology and debitage features (e.g., typology, cutting edge length, platform dimensions) may yet reveal a more reliable signal of knapping quality. Results of the Quality model also seem to suffer from the uneven distribution and discrete rather than continuous nature of scores on our RPM and n-Back tests. We could address concerns about sampling variation on these dimensions with larger samples or by pre-screening participants to ensure more even representation. Alternative psychometric tests (e.g., complete rather than the short version of the RPM) might also provide more sensitive and continuous measures.

Another major limitation our study shares with all other published experiments on knapping skill acquisition is that we do not address variation in social and cultural context or teaching style [@kline2015] . Currently, we have little basis other than personal experience/tradition [@callahan1979; @whittaker1994; @shea2015] and theoretical speculation [@stout2013; @whiten2015] from which to assess which pedagogical techniques are most effective even in WEIRD contexts. No study to date has considered how variation in teacher skill [@shea2015] or social relationship to participants might impact learning under different conditions. Addressing these questions would require a significant research program, including cross-cultural comparative studies [@barrett2020; @fuentes2016] and more naturalistic study designs. While costly, such research would produce results of broad relevance to anthropologists, biologists, psychologists, and sociologists interested in teaching and learning.

# Conclusions

Our exploratory study produced findings with both theoretical and methodological implications. Perhaps the clearest result is that the effects of different experimental learning conditions and their interaction with individual cognitive and motor differences are complicated, and researchers must interpret them with care. For example, we did not find robust effects of instruction on mean group performance. Still, we did find that teaching reduced variation knapping rate, methods, and outcomes, thus providing a "safety net" against failed transmission. This increase in reliability of skill reproduction offers further support for the hypothetical co-evolution of teaching, language, and tool-making [@morgan2015;@stout2010]. At the same time, we found that two hours of practice was insufficient to document learning-related performance change. This result implies that we may need longer experiment durations to study learning, even for basic Oldowan-like flaking. Researchers should take care when interpreting results from shorter studies. It is thus important to have a quantitative standard of proficiency, ideally archaeological [@pargeter2019;@stout2019] but at least experimental (this study), against which to compare performance. Here, it was clear that participants remained in an early learning stage far from expert performance and that teaching had important effects on how learners explored the novel task space. Instruction allowed learners to maximize both quality and quantity rather than trading off between them and dramatically changed the effects of individual differences in strength, visuospatial working memory, and social learning tendencies on knapping outcomes. These results highlight the value of including individual difference measures in knapping experiments and suggest that the presence/absence of teaching in the Paleolithic would have fundamentally altered associated selective pressures [@fuentes2015; @stout2019] in addition to affecting the reliability of skill reproduction. To unravel these complicated and nuanced relationships, future experiments might benefit from more extended instruction periods, video-based quantitative analyses of knapping behaviors [@geribàs2010; @stout2021], more refined psychometric and lithic measures, a greater diversity of social and cultural contexts [@barrett2020], and explicit attention to teacher skill, style, and interpersonal relationships with participants [@bevilacqua2019].

# References